Machine Learning, Architectural Styles and Property Values

### 2020-04-14 text/manuscript_jrefe.tex
TeX file generated

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Machine Learning, Architectural Styles and Property Values},
            pdfauthor={Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson (University of Alabama)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Machine Learning, Architectural Styles and Property Values}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{13 April 2020}

\usepackage{soul}
\usepackage{color}
\sethlcolor{yellow}
\usepackage{float}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{multirow}
\doublespacing
\usepackage[backend=bibtex, style=authoryear]{biblatex}

@@

If new developments were only pleasing to the eye, so Britain's
\emph{Building Better, Building Beautiful Commission} advising
Government, nimbyism would cease and housing supply could finally reach
the levels demanded by a growing and more affluent population (The
Economist 2018). The inaugural chair of this new commission boldly
suggested that if we built \emph{``as our Georgian and Victorian
forebears built} {[}\ldots{}{]}. \emph{All objections to new building
would slip away in the sheer relief of the public''} (Scruton 2018).
Even Prince Charles, in similar spirit, put forward ten principles for
urban growth and architecture that emphasize tradition and aesthetics
(HRH The Prince of Wales 2014). When governments and princes occupy
themselves with beauty, economists surely may take a closer look as
well. In this paper, we aim to identify the architectural styles of
residential buildings using computer vision techniques and to
empirically search for any transaction price premia associated with
these styles. If aesthetics were as strong a force in the built
environment as claimed, we should indeed find a significant effect on
property prices.

This paper assesses the value of architectural aesthetics head on by
classifying the exterior styles of residential buildings using street
level images. Recent work illustrates how street level imagery and
machine learning classification is an efficient and powerful combination
for measuring previously unobserved characteristics of the urban
environment. Naik et al. (2017) describe how neighborhood demographics
may impact the physical appearance of neighborhoods. Gebru et al. (2017)
use classified vehicle make and model information to predict income,
race, education, and voting patterns at the precinct level. Glaeser et
al. (2018) predict income in New York City. Naik, Raskar, and Hidalgo
(2016) create a neighborhood safety based Streetscore which is shown to
be highly correlated with neighborhood population density and household
income. De Nadai et al. (2016) find that greenery and street facing
windows contribute to a positive appearance of safety while Liu et al.
(2017) evaluate the quality and upkeep of the built environment along
Beijing's streets.

Glaeser, Kincaid, and Naik (2018) push the level of observation from the
block, street, or street-section level to the individual
\emph{building level}. Utilizing images of buildings' exteriors
collected from Google Street
View\footnote{\href{https://www.google.co.uk/maps}{https://www.google.co.uk/maps}},
and to a lesser degree interior images from Zillow, they find that looks
matter, at least in Boston: A one standard deviation improvement of a
building's exterior is associated with an additional USD 70,000 in home
value. Intuitively, the link between good looks and value is
bi-directional: The appearance of buildings that went through
foreclosure deteriorated significantly (Glaeser, Kincaid, and Naik
2018).

Zooming in at individual buildings significantly increases the benefits
of using mass collected street level imagery in economic research:
property characteristics previously deemed ``unobservable'' can be
directly observed in an accurate, objective, automatic and therefore
cost effective way. Deriving additional variables from unconventional
data sources like 3D airborne laser scanning (or in our case Google
Street View) is fundamental since this provides ``essential determinants
influencing real estate prices {[}which{]} are constantly missing and
are not accessible in official and mass appraiser databases'' (Helbich
et al. 2013).

@@

The aesthetics of different architectural styles make for an easy
cocktail party conversation topic -- and they may also influence market
prices. Using architectural assessments by human experts, Buitelaar and
Schilder (2017) find a sizable premium of 5\% for new buildings in the
Netherlands that refer to traditional styles and a staggering 15\%
premium for new buildings that closely follow traditional shapes, facade
composition and details. Their study carefully disentangles the
architectural style from other unobserved characteristics such as
building quality, differences in location, or year of construction.
These controls are crucial, as earlier work has established that age and
style variables tend to be highly correlated. Coulson and McMillen
(2008), for instance, suggest a non-parametric estimator and establish a
U-shaped age function and distinct price discounts for postwar and
contemporary styles (vis-à-vis more historic styles). Francke and van de
Minne (2017) investigate the depreciation of residential real estate in
the Netherlands and decompose land versus structure values singling out
the effect of ``physical deterioration, functional obsolescence and and
vintage effects''. They find that buildings from the 1930s carry a
strong price premium.

A large scale assessment of buildings' exteriors will allow for an
analysis of the externalities of architecture. Buildings hardly ever
stand in isolation and Ahlfeldt and Mastro (2012) investigate the
influence a building's architecture exerts on its surroundings. They
observe a positive price effect for residential buildings in the direct
proximity of iconic homes by Frank Lloyd Wright in Oak Park, Illinois. A
building's exterior does not need to be an architectural masterpiece to
co-determine the value of other houses close by. Homogeneity of building
shapes within street segments does influence property values. A
similarly shaped neighboring building is value enhancing while proximity
to a wildly different neighboring shape, everything else remaining
equal, is detrimental to property values (Lindenthal 2017a).

Unfortunately, the traditional approaches chosen by e.g.~Buitelaar and
Schilder (2017) or Ahlfeldt and Mastro (2012) do not scale well. Each
observation has to be classified into architectural styles by a human
expert, which is time-consuming and costly, imposing an upper limit on
the number of observations and the level of detail captured for each
observation in any given sample. Google Street View offers a solution to
this nexus as it captures the images of almost all buildings in many
cities around the world at high level of accuracy and detail. The
challenge we address in this paper is to extract and utilize building
level information from this ubiquitous sensor, using freely available
deep learning techniques.

@@

Previous work sourcing imagery from Google Street View has mostly
focused on neighborhood or precinct characteristics where exact spatial
assignment of objects is not required and street-level images suffice.
For major US cities, the accuracy of building image search results is
higher, which should help studies such as Glaeser, Kincaid, and Naik
(2018).

@@

By combining land registry maps with the snapshot location, we are able
to automatically build a set of building frontage images for
approximately 48,000 properties in Cambridge (UK). The Ordnance Survey
maps do not provide any use type classifications so we collect images on
all buildings that have a ground plate between 40 and 250 \(m^2\). This
method, however, could easily be adapted to photograph buildings in
locations for which high-resolution building level maps are not
available by using LIDAR based building outlines instead, which would be
in the spirit of Glaeser et al. (2018).

@@

We download all pictures at the highest resolution offered by Google's
Street View
API\footnote{\href{https://developers.google.com/maps/documentation/streetview/intro}{https://developers.google.com/maps/documentation/streetview/intro}}
(640x640 color pixels). We then obtain 2048-dimensional feature vectors
for each picture, using the \emph{Inception-v3} deep convolutional
neural network (Szegedy et al. 2015). \emph{Inception-v3} has been
trained for the ImageNet Large Scale Visual Recognition Challenge
(ILSVRC)\footnote{\url{http://image-net.org/challenges/LSVRC/}}, which
evaluates image classification and object detection algorithms for a
wide range of objects. The pretrained classifications would allow us to
identify pets, vehicles or people on the pictures -- assessing
architectural style, however, is beyond the canned classifiers'
capabilities.

Glaeser, Kincaid, and Naik (2018) rely on an different ILSVRC
competitor, \emph{Resnet-101} (He et al. 2016). They reduce the
extracted feature vectors to lower dimensionality (1024 to 100
dimensions) based on principal component analysis (PCA). We follow a
different strategy and double the dimensionality by including the
feature vector of the closest building in the sample, which, in most
cases, is the direct neighbor. The feature vectors of neighbors have
been collected in exactly the same way as those for other buildings in
the sample. Doubling up allows us to model spatial dependencies in
building styles, similar to spatially correlated land cover
classifications in Ghimire, Rogan, and Miller (2010).

@@

Will the resulting classifiers be biased with respect to property
values, picking up correlated but irrelevant cues such as greenery,
upkeep, or cars brands instead of building exteriors? Are, for instance,
more valuable houses more likely to be classified as the popular
\emph{Late Victorian} style while starter homes are disproportionally
more likely to be classified as e.g.~\emph{Early Victorian}? We expect
that in cases in which the ensemble provided a high confidence score,
building features are clearly detectable on the pictures. For close
calls, however, where models within the ensemble cannot agree, other
factors might tilt the scale towards an incorrect class. A sizable
difference in estimated marginal prices for machine's classifications
versus the estimates for architects' classifications in the subsequent
hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

@@

Residential real estate transactions are public data in the UK,
collected and published by the Land Registry (Land Registry 2017). The
records include the date of transaction, price paid, street address, a
classification of the property type (flat, detached, semi-detached, or
terraced house), the estate type (freehold or leasehold) and an
indicator for newly built properties. We select transactions from
Cambridge proper which were recorded between January 1995 and October
2018, excluding any leaseholds, flats and properties classified as type
``other'', and sales with prices below £50,000 or in excess of
£2,000,000. Linking recently taken images to sales data from 24 years
assumes that houses have not changed their architectural styles -- which
seems reasonable. We are not aware of any conversions of just the
exterior of buildings. In case of full redevelopments, we exclude any
sales for a given address pre-dating the redevelopment, ensuring that a
currently taken pictures represents the house at all included
transactions. Table \ref{tab:sumstats} presents summary statistics for
the sample.

@@

The Ordnance Survey \emph{AddressBase} (Ordnance Survey 2017a) links
street addresses to building outlines on Ordnance Survey maps (Ordnance
Survey 2017b), which allows us to calculate the buildings floor plate
(in \(m^2\)) and to estimate the building's volume from digital
elevation models (Environment Agency 2015), as suggested by Lindenthal
(2017b). We control for the location of each building by calculating the
distance to the city center proxied by Great St.~Mary's Church, and
non-parametrically by using 69 indicator variables for each of the
smallest census tracts (\emph{Lower Super Output Areas}, LSOA)
subdividing Cambridge (Office for National Statistics 2019). LSOAs
typically have 1,000--3,000 residents and 400--1,200 households of
comparable economic and socio-demographic characteristics (Office for
National Statistics 2017).

@@

Merging the machine's classification with sales data (Land Registry
2017) we estimate a hedonic regression equation that establishes
marginal prices for the building style (similar to Moorhouse and Smith
1994; Asabere, Hachey, and Grubaugh 1989; Vandell and Lane 1989; Fuerst,
McAllister, and Murray 2011; Plaut and Uzulena 2006), among other
characteristics: \begin{equation} \label{eq:hed}
  ln(Price_i) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Style}_i + \eta\mathbf{StyleNeigh}_i + \iota\mathbf{Style\cdot StyleNeigh}_i  +\gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

Here, the natural logarithm of sales prices is explained by a linear
combination of hedonic attributes described in vector \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's estimated \(\mathbf{Style}\) and the
prevailing styles of other buildings in the direct proximity
\((\mathbf{StyleNeigh})\). \(\mathbf{Style\cdot StyleNeigh}\) is a
vector of interaction terms for the building's and the neighborhood's
dominant style. The intercept is denoted by \(\alpha\) while
\(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\eta}\), \(\bm{\iota}\),
\(\bm{\gamma}\) and \(\bm{\lambda}\) are vectors of regression
coefficients. \(\epsilon\) is the IID error term. Heteroscedastic robust
standard errors will be reported.

Are buildings with different appearances imperfect substitutes catering
to multiple groups of households with distinct style or vintage
preferences? Buitelaar and Schilder (2017) indicate that any premium for
an architectural style must stem from either differences in construction
prices (which they do not find in their Dutch sample) or from supply
constraints, as new construction potentially does not capture the demand
for traditional styles. For Cambridge, new supply will inevitably be of
either contemporary or revival style as historic vintages (and their
styles), by definition, are not supplied any more. Estimating Eq.
\ref{eq:hed} for a subset of newly constructed buildings will show
whether construction prices or supply constraints for new homes built
according to different architectural styles persist -- if too few
vernacular buildings were built, prices should reflect such a shortage.
Singling out new buildings crucially controls for the otherwise
unobserved age of buildings, which is tightly intertwined with its
aesthetics.

@@

A visual inspection of classifications is additionally reassuring.
Figure \ref{fig:vint_examples} displays the pictures which carry the
highest out-of-sample scores for each style The model is clearly able to
differentiate based on small cues, even when only parts of the facade
are captured on the picture.

@@

The majority of misclassification is concentrated in temporally adjacent
eras: 10 percent of \emph{Late Victorian/Edwardian} buildings are
labelled as \emph{Early Victorian} or 18 percent of \emph{Postwar}
buildings are erroneously regarded as stemming from the \emph{Interwar}
period. Hardly ever is a contemporary building mistaken for a historic
home.

@@

The sample images classified by architects is not necessarily
representative of the overall population so we re-estimate the model
using machine classifications for \emph{all} buildings for which we were
able to extract usable pictures from Google Street View. Overall, the
style coefficients become more positive relative to the base, implying
that our automatic classification is biased and tends to classify
unattractive buildings to the \emph{Contemporary} style. Reversely, more
attractive buildings are disproportionally often misclassified as
\emph{Late Vic./Edw.}, pushing up the corresponding coefficient to
\(0.11\). This bias vanishes in the third model, which uses only
confidently estimated machine styles (Column 3).

Models 4--6 follow the same logic as 1--3 but are estimated for
\emph{new} buildings only. Basically this contrasts houses which might
\emph{look} different from the street, having either contemporary or
revival facades, but which are all modern homes at their core.
Differences in e.g.~materials, floor plans, green space and gardens are
minimal. For newly built homes, no effect of \emph{Revival} architecture
on price can be found for high confidence ML estimates, neither for
classifications by machines nor by architects. After controlling for
location, building characteristics and quality, buyers show no
willingness to pay a premium for \emph{Revival} architecture (Columns 4
and 6).

Introducing a variable capturing the most prominent style of buildings
on the same street and within 100 m (\(\mathbf{VintNeigh}\)) does not
change the pecking order of building styles much (Table
\ref{tab:hedreg}, Column 7). Neighboring styles are compared to the base
case of a \emph{Contemporary} house in a \emph{Contemporary} street. The
negative coefficients on dissimilar neighboring styles support a finding
by Lindenthal (2017b) that shows that a harmonious match of a building's
shape with its direct environment leads to a price premium -- or a
discount in case of shape mismatches. In relative terms, \emph{Postwar}
buildings are the least popular neighbors. The difference between
\emph{Revival} and \emph{Contemporary} neighbors is not statistically
significant. We fail to find evidence in favor of positive externalities
from revival architecture. Everything else equal, buildings within
ensembles of \emph{Contemporary} generally do not achieve higher
transaction prices than buildings in more historic areas (as Table
\ref{tab:comb} will confirm later).

@@

\hypertarget{bibliography}{%
\section*{Bibliography}\label{bibliography}}
\addcontentsline{toc}{section}{Bibliography}

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Machine Learning, Architectural Styles and Property Values},
            pdfauthor={Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson (University of Alabama)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{soul}
\usepackage{color}
\sethlcolor{yellow}
\usepackage{float}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{multirow}
\doublespacing
\usepackage[backend=bibtex, style=authoryear]{biblatex}
\usepackage[]{biblatex}
\addbibresource{references.bib}

\title{Machine Learning, Architectural Styles and Property Values}
\author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
\date{14 April 2020}

@@

If new developments were only pleasing to the eye, so Britain's
\emph{Building Better, Building Beautiful Commission} advising
Government, nimbyism would cease and housing supply could finally reach
the levels demanded by a growing and more affluent population
\autocite{economistlineofbeauty2018}. The inaugural chair of this new
commission boldly suggested that if we built \emph{``as our Georgian and
Victorian forebears built} {[}\ldots{}{]}. \emph{All objections to new
building would slip away in the sheer relief of the public''}
\autocite{scruton2018}. Even Prince Charles, in similar spirit, put
forward ten principles for urban growth and architecture that emphasize
tradition and aesthetics \autocite{princecharles2014}. When governments
and princes occupy themselves with beauty, economists surely may take a
closer look as well. In this paper, we aim to identify the architectural
styles of residential buildings using computer vision techniques and to
empirically search for any transaction price premia associated with
these styles. If aesthetics were as strong a force in the built
environment as claimed, we should indeed find a significant effect on
property prices.

This paper assesses the value of architectural aesthetics head on by
classifying the exterior styles of residential buildings using street
level images. Recent work illustrates how street level imagery and
machine learning classification is an efficient and powerful combination
for measuring previously unobserved characteristics of the urban
environment. \textcite{Naik2017} describe how neighborhood demographics
may impact the physical appearance of neighborhoods.
\textcite{Gebru2017} use classified vehicle make and model information
to predict income, race, education, and voting patterns at the precinct
level. \textcite{Glaeser2016} predict income in New York City.
\textcite{Naik2016AER} create a neighborhood safety based Streetscore
which is shown to be highly correlated with neighborhood population
density and household income. \textcite{Nadai2016} find that greenery
and street facing windows contribute to a positive appearance of safety
while \textcite{Liu2017} evaluate the quality and upkeep of the built
environment along Beijing's streets.

\textcite{GlaeserKincaidNaik2018} push the level of observation from the
block, street, or street-section level to the individual
\emph{building level}. Utilizing images of buildings' exteriors
collected from Google Street
View\footnote{\href{https://www.google.co.uk/maps}{https://www.google.co.uk/maps}},
and to a lesser degree interior images from Zillow, they find that looks
matter, at least in Boston: A one standard deviation improvement of a
building's exterior is associated with an additional USD 70,000 in home
value. Intuitively, the link between good looks and value is
bi-directional: The appearance of buildings that went through
foreclosure deteriorated significantly
\autocite{GlaeserKincaidNaik2018}.

Zooming in at individual buildings significantly increases the benefits
of using mass collected street level imagery in economic research:
property characteristics previously deemed ``unobservable'' can be
directly observed in an accurate, objective, automatic and therefore
cost effective way. Deriving additional variables from unconventional
data sources like 3D airborne laser scanning (or in our case Google
Street View) is fundamental since this provides ``essential determinants
influencing real estate prices {[}which{]} are constantly missing and
are not accessible in official and mass appraiser databases''
\autocite{Helbich2013}.

@@

The aesthetics of different architectural styles make for an easy
cocktail party conversation topic -- and they may also influence market
prices. Using architectural assessments by human experts,
\textcite{Buitelaar2017} find a sizable premium of 5\% for new buildings
in the Netherlands that refer to traditional styles and a staggering
15\% premium for new buildings that closely follow traditional shapes,
facade composition and details. Their study carefully disentangles the
architectural style from other unobserved characteristics such as
building quality, differences in location, or year of construction.
These controls are crucial, as earlier work has established that age and
style variables tend to be highly correlated. \textcite{Coulson2008},
for instance, suggest a non-parametric estimator and establish a
U-shaped age function and distinct price discounts for postwar and
contemporary styles (vis-à-vis more historic styles).
\textcite{Francke2017a} investigate the depreciation of residential real
estate in the Netherlands and decompose land versus structure values
singling out the effect of ``physical deterioration, functional
obsolescence and and vintage effects''. They find that buildings from
the 1930s carry a strong price premium.

A large scale assessment of buildings' exteriors will allow for an
analysis of the externalities of architecture. Buildings hardly ever
stand in isolation and \textcite{Ahlfeldt2012} investigate the influence
a building's architecture exerts on its surroundings. They observe a
positive price effect for residential buildings in the direct proximity
of iconic homes by Frank Lloyd Wright in Oak Park, Illinois. A
building's exterior does not need to be an architectural masterpiece to
co-determine the value of other houses close by. Homogeneity of building
shapes within street segments does influence property values. A
similarly shaped neighboring building is value enhancing while proximity
to a wildly different neighboring shape, everything else remaining
equal, is detrimental to property values \autocite{Lindenthal2017b}.

Unfortunately, the traditional approaches chosen by
e.g.~\textcite{Buitelaar2017} or \textcite{Ahlfeldt2012} do not scale
well. Each observation has to be classified into architectural styles by
a human expert, which is time-consuming and costly, imposing an upper
limit on the number of observations and the level of detail captured for
each observation in any given sample. Google Street View offers a
solution to this nexus as it captures the images of almost all buildings
in many cities around the world at high level of accuracy and detail.
The challenge we address in this paper is to extract and utilize
building level information from this ubiquitous sensor, using freely
available deep learning techniques.

@@

Previous work sourcing imagery from Google Street View has mostly
focused on neighborhood or precinct characteristics where exact spatial
assignment of objects is not required and street-level images suffice.
For major US cities, the accuracy of building image search results is
higher, which should help studies such as
\textcite{GlaeserKincaidNaik2018}.

@@

By combining land registry maps with the snapshot location, we are able
to automatically build a set of building frontage images for
approximately 48,000 properties in Cambridge (UK). The Ordnance Survey
maps do not provide any use type classifications so we collect images on
all buildings that have a ground plate between 40 and 250 \(m^2\). This
method, however, could easily be adapted to photograph buildings in
locations for which high-resolution building level maps are not
available by using LIDAR based building outlines instead, which would be
in the spirit of \textcite{Glaeser2016}.

@@

We download all pictures at the highest resolution offered by Google's
Street View
API\footnote{\href{https://developers.google.com/maps/documentation/streetview/intro}{https://developers.google.com/maps/documentation/streetview/intro}}
(640x640 color pixels). We then obtain 2048-dimensional feature vectors
for each picture, using the \emph{Inception-v3} deep convolutional
neural network \autocite{Szegedy2015}. \emph{Inception-v3} has been
trained for the ImageNet Large Scale Visual Recognition Challenge
(ILSVRC)\footnote{\url{http://image-net.org/challenges/LSVRC/}}, which
evaluates image classification and object detection algorithms for a
wide range of objects. The pretrained classifications would allow us to
identify pets, vehicles or people on the pictures -- assessing
architectural style, however, is beyond the canned classifiers'
capabilities.

\textcite{GlaeserKincaidNaik2018} rely on an different ILSVRC
competitor, \emph{Resnet-101} \autocite{He2016}. They reduce the
extracted feature vectors to lower dimensionality (1024 to 100
dimensions) based on principal component analysis (PCA). We follow a
different strategy and double the dimensionality by including the
feature vector of the closest building in the sample, which, in most
cases, is the direct neighbor. The feature vectors of neighbors have
been collected in exactly the same way as those for other buildings in
the sample. Doubling up allows us to model spatial dependencies in
building styles, similar to spatially correlated land cover
classifications in \textcite{Ghimire2010a}.

@@

Will the resulting classifiers be biased with respect to property
values, picking up correlated but irrelevant cues such as greenery,
upkeep, or cars brands instead of building exteriors? Are, for instance,
more valuable houses more likely to be classified as the popular
\emph{Late Victorian} style while starter homes are disproportionally
more likely to be classified as e.g.~\emph{Early Victorian}? We expect
that in cases in which the ensemble provided a high confidence score,
building features are clearly detectable on the pictures. For close
calls, however, where models within the ensemble cannot agree, other
factors might tilt the scale towards an incorrect class. A sizable
difference in estimated marginal prices for machine classifications
versus the estimates for architects' classifications in the subsequent
hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

@@

Residential real estate transactions are public data in the UK,
collected and published by the Land Registry
\autocite{LandRegistry2016a}. The records include the date of
transaction, price paid, street address, a classification of the
property type (flat, detached, semi-detached, or terraced house), the
estate type (freehold or leasehold) and an indicator for newly built
properties. We select transactions from Cambridge which were recorded
between January 1995 and October 2018, excluding any leaseholds, flats
and properties classified as type ``other'', and sales with prices below
£50,000 or in excess of £2,000,000. Linking recently taken images to
sales data from 24 years assumes that houses have not changed their
architectural styles -- which seems reasonable. We are not aware of any
conversions of just the exterior of buildings. In case of full
redevelopments, we exclude any sales for a given address pre-dating the
redevelopment, ensuring that a currently taken picture represents the
house at all included transactions. Table \ref{tab:sumstats} presents
summary statistics for the sample.

@@

The Ordnance Survey \emph{AddressBase} \autocite{OrdnanceSurvey2017}
links street addresses to building outlines on Ordnance Survey maps
\autocite{OrdnanceSurvey2017a}, which allows us to calculate the
buildings floor plate (in \(m^2\)) and to estimate the building's volume
from digital elevation models \autocite{EnvironmentAgency2015}, as
suggested by \textcite{Lindenthal2017}. We control for the location of
each building by calculating the distance to the city center proxied by
Great St.~Mary's Church, and non-parametrically by using 69 indicator
variables for each of the smallest census tracts (\emph{Lower Super
Output Areas}, LSOA) subdividing Cambridge \autocite{ONS2017}. LSOAs
typically have 1,000--3,000 residents and 400--1,200 households of
comparable economic and socio-demographic characteristics
\autocite{ONS2017b}.

@@

Merging the machine's classification with sales data
\autocite{LandRegistry2016a} we estimate a hedonic regression equation
that establishes marginal prices for the building style
\autocites[similar
to][]{Moorhouse1994}{Asabere1989}{Vandell1989}{Fuerst2011}{Plaut2006},
among other characteristics: \begin{equation} \label{eq:hed}
  ln(Price_i) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Style}_i + \eta\mathbf{StyleNeigh}_i + \iota\mathbf{Style\cdot StyleNeigh}_i  +\gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

Here, the natural logarithm of sales prices is explained by a linear
combination of hedonic attributes described in vector \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's estimated \(\mathbf{Style}\) and the
prevailing styles of other buildings in the direct proximity
\((\mathbf{StyleNeigh})\). \(\mathbf{Style\cdot StyleNeigh}\) is a
vector of interaction terms for the building's and the neighborhood's
dominant style. The intercept is denoted by \(\alpha\) while
\(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\eta}\), \(\bm{\iota}\),
\(\bm{\gamma}\) and \(\bm{\lambda}\) are vectors of regression
coefficients. \(\epsilon\) is the IID error term. Heteroscedasticity
robust standard errors will be reported.

Are buildings with different appearances imperfect substitutes catering
to multiple groups of households with distinct style or vintage
preferences? \textcite{Buitelaar2017} indicate that any premium for an
architectural style must stem from either differences in construction
prices (which they do not find in their Dutch sample) or from supply
constraints, as new construction potentially does not capture the demand
for traditional styles. For Cambridge, new supply will inevitably be of
either contemporary or revival style as historic vintages (and their
styles), by definition, are not supplied any more. Estimating Eq.
\ref{eq:hed} for a subset of newly constructed buildings will show
whether construction prices or supply constraints for new homes built
according to different architectural styles prevail -- if too few
vernacular buildings were built, prices should reflect such a shortage.
Singling out new buildings crucially controls for the otherwise
unobserved age of these buildings, which is tightly intertwined with
their aesthetics.

@@

A visual inspection of classifications is additionally reassuring.
Figure \ref{fig:vint_examples} displays the pictures which carry the
highest out-of-sample scores for each style. The model is clearly able
to differentiate based on small cues, even when only parts of the facade
are captured on the picture.

@@

The majority of misclassifications is concentrated in temporally
adjacent eras: 10 percent of \emph{Late Victorian/Edwardian} buildings
are labelled as \emph{Early Victorian} or 18 percent of \emph{Postwar}
buildings are erroneously regarded as stemming from the \emph{Interwar}
period. Hardly ever is a contemporary building mistaken for a historic
home.

@@

The sample of images classified by architects is not necessarily
representative of the overall population so we re-estimate the model
using machine classifications for \emph{all} buildings for which we were
able to extract usable pictures from Google Street View. Overall, the
style coefficients become more positive relative to the base, implying
that our automatic classification is biased and tends to classify bland
or low quality buildings to the \emph{Contemporary} style. Reversely,
more attractive or higher quality buildings are disproportionately often
misclassified as \emph{Late Vic./Edw.}, pushing up the corresponding
coefficient to \(0.11\). This bias vanishes in the third model, which
uses only confidently estimated machine styles (Column 3).

Models 4--6 follow the same logic as 1--3 but are estimated for
\emph{new} buildings only, which are either in the contemporary or the
revival style. Basically this contrasts houses which might \emph{look}
different from the street, having either contemporary or revival
facades, but which are all modern homes at their core. Differences in
e.g.~materials, floor plans, green space and gardens are minimal. For
newly built homes, no effect of \emph{Revival} architecture on price can
be found for high confidence ML estimates, neither for classifications
by machines nor by architects. After controlling for location, building
characteristics and quality, buyers show no willingness to pay a premium
for \emph{Revival} architecture (Columns 4 and 6).

Introducing a variable capturing the most prominent style of buildings
on the same street and within 100 m (\(\mathbf{StyleNeigh}\)) does not
change the pecking order of building styles much (Table
\ref{tab:hedreg}, Column 7). Neighboring styles are compared to the base
case of a \emph{Contemporary} house in a \emph{Contemporary} street. The
negative coefficients on dissimilar neighboring styles support a finding
by \textcite{Lindenthal2017} that shows that a harmonious match of a
building's shape with its direct environment leads to a price premium --
or a discount in case of shape mismatches. In relative terms,
\emph{Postwar} buildings are the least popular neighbors. The difference
between \emph{Revival} and \emph{Contemporary} neighbors is not
statistically significant. We fail to find evidence in favor of positive
externalities from revival architecture. Everything else equal,
buildings within ensembles of \emph{Contemporary} generally do not
achieve higher transaction prices than buildings in more historic areas
(as Table \ref{tab:comb} will confirm later).

@@

\printbibliography[title=Bibliography]
